Cognate beginnings to bilingual lexical acquisition

### 2023-05-25 docs/index.qmd
Make site

---
title: "Trajectories"
subtitle: "Lab notes"
date: "Updated: `r format(Sys.Date(), '%Y/%m/%d')`"
authors:
  - name: Gonzalo Garcia-Castro
    orcid: 0000-0002-8553-4209
    email: gonzalo.garciadecastro@upf.edu
    corresponding: true
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
        address: Ramon Trias fargas 25-27
        city: Barcelona
        state: Spain
        postal-code: 08005
  - name: Daniela S. Ávila-Varela
    orcid: 0000-0002-3518-8117
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
  - name: Ignacio Castillejo
    orcid: 0000-0001-7445-0416
    affiliations:
      - name: Universidad Autónoma de Madrid
        department: Departamento de Psicología
  - name: Nuria Sebastian-Galles
    orcid: 0000-0001-6938-2498
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies. Floccia et al. (2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch). The mechanisms underlying this effect remain unknown. We explore two compatible scenarios. First, we test whether cognates are learnt earlier than non-cognates. This would account for the difference in vocabulary size associated to the amount of shared cognates across languages. Second, we explore the possibility that the word-forms of one language interact with those in the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap. This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals. We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
format:
    html:
        code-copy: hover
        reference-location: margin
        code-overflow: scroll
link-external-newwindow: true
code-fold: true
code-line-numbers: true
fig-dpi: 1000
page-layout: full
warning: false
echo: false
toc: true
colorlinks: true
number-sections: true
csl: "../assets/apa7.csl"
bibliography: "../assets/references.bib"
---

```{r}
#| label: setup
#| include: false

# import packages
library(conflicted)
library(ggplot2)
library(patchwork)
library(dplyr)
library(tidyr)
library(purrr)
library(gt)
library(forcats)
library(stringr)
library(tibble)
library(tidybayes)
library(janitor)
library(glue)
library(brms)
library(scales)
library(marginaleffects)
library(lubridate)
library(bvqdev)

# load targets as R objects
tar_load(participants)
tar_load(responses)
tar_load(items)
tar_load(bvq_data)
tar_load(model_fit_0)
tar_load(model_fit_1)
tar_load(model_fit_2)
tar_load(model_fit_3)
tar_load(model_fit_4)
tar_load(model_fit_4_prior)
tar_load(model_fit_prior)
tar_load(model_loos)
tar_load(posterior_draws)
tar_load(posterior_draws_re)
tar_load(marginal_effects_epreds)

# resolve namespace conflicts
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

# set ggplot theme and colour palette
theme_set(theme_custom()) # set custom ggplot theme
clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

options(ggplot2.ordinal.fill = clrs,
        ggplot2.ordinal.colour = clrs,
        ggplot2.discrete.fill = clrs,
        ggplot2.discrete.colour = clrs,
        ggplot2.continuous.fill = ggplot2::scale_color_gradient,
        ggplot2.continuous.colour = ggplot2::scale_color_gradient)

rope_interval <- c(lower = -0.1, upper = +0.1)
```

@@

```{r}
#| label: questionnaire
#| echo: false
#| message: false
#| warning: false
data("pool", package = "bvqdev")
n_categories <- distinct(pool, semantic_category) %>% nrow()
n_items <- length(unique(pool$item))
n_item_language <- count(pool, language) %>% 
    pull(n) %>%
    set_names(c("catalan", "spanish"))
```

@@

```{r apatb-participants}
#| label: tab-participants
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan. Information about language degree of exposure (DoE) was provided by families before filling the questionnaire. In this table, a 100\\% indicates that the participant was exclusively exposed to Catalan, and never to Spanish. A 0\\% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish. A 50\\% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each. The 100\\%, 0\\%, and 50\\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively. For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4 month-wide bins. Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
participants |>
    mutate_at(vars(starts_with("doe_")), 
              (\(x) floor(x * 10) / 10)) |> 
    mutate(
        age = cut(floor(age),
                  c(10, 14, 18, 22, 26, 30, 34, 36),
                  include.lowest = TRUE),
        doe_catalan = cut(
            doe_catalan,
            c(0, 0.25, 0.5, 0.75, 1),
            labels = c("0-25%",
                       "25-50%", 
                       "50-75%",
                       "75-100%"),
            include.lowest = TRUE
        ) |>
            as.character()
    ) |>
    count(age, doe_catalan) |>
    arrange(desc(doe_catalan)) |>
    pivot_wider(names_from = age, values_from = n) |>
    gt() |>
    summary_rows(
        columns = 2:7,
        fns = list("Total" = ~ sum(., na.rm = TRUE)),
        formatter = fmt_integer
    ) |>
    sub_missing(everything(), everything(), missing_text = "--") |>
    cols_label(doe_catalan = "Catalan exposure") |>
    tab_spanner("Age (months)", 2:8) |>
    tab_style(cell_text(align = "left"),
              list(cells_body(doe_catalan),
                   cells_column_labels(doe_catalan)))
```

@@

```{r}
#| label: tbl-participants-edu
#|  tbl-cap: "Participant exposure to the language (DoE) and socio-economic status (SES)."
participants %>%
    mutate(id = as.character(id)) |> 
    left_join(bvq_data$logs) %>%
    mutate(
        edu_parent = edu_parent |> 
            as.factor() |> 
            fct_na_value_to_level(),
        doe_catalan = cut(doe_catalan,
                          breaks = seq(0, 1, 0.2),
                          include.lowest = TRUE)
    ) %>%
    count(doe_catalan, edu_parent) %>%
    right_join(
        expand_grid(
            doe_catalan = levels(.$doe_catalan),
            edu_parent = unique(bvq_data$logs$edu_parent)
        )
    ) %>%
    rename(total = n) %>%
    pivot_wider(names_from = edu_parent,
                values_from = total,
                values_fill = 0) %>%
    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>%
    rename(Total = total) %>%
    select(-"NA") %>%
    gt() |> 
    summary_rows(columns = Complementary:Primary,
                 fns = list(Total = ~sum(., na.rm = TRUE)),
                 formatter = fmt_integer) %>%
    cols_label(doe_catalan = "DoE Catalan") %>%
    sub_missing(columns = everything(),
                missing_text = "--") %>%
    tab_spanner("Educational attainment", 
                Secondary:Primary) %>%
    tab_style(
        style = list(cell_borders(weight = px(2),
                                  sides = "left", 
                                  color = "grey50")),
        locations = cells_body(columns = 9,
                               rows = 1:5)
    ) %>%
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(columns = 2:7)
    ) %>%
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = c(1, 8))
    ) %>%
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(spanners = "Educational attainment")
    )
```

@@

items %>%
    unnest(list) %>%
    summarise(n = n(),
              across(c(freq, n_phon, lv),
                     lst(mean, sd, min, max)),
              .by = c(language, list)) |> 
    gt(groupname_col = "language", 
       rowname_col = "list") %>%
    fmt_number(-c(list, n),
               decimals = 1) %>%
    tab_spanner("Frequency (Zipf)",
                columns = starts_with("freq_")) %>%
    tab_spanner("Length (Phonemes)",
                columns = starts_with("n_phon_")) %>%
    tab_spanner("Levenshtein (cognateness)", 
                columns = starts_with("lv_")) %>%
    cols_merge_uncert(col_val = freq_mean,
                      col_uncert = freq_sd) %>%
    cols_merge_uncert(col_val = n_phon_mean,
                      col_uncert = n_phon_sd) %>%
    cols_merge_uncert(col_val = lv_mean, 
                      col_uncert = lv_sd) %>%
    cols_merge_range(col_begin = freq_min, 
                     col_end = freq_max) %>%
    cols_merge_range(col_begin = n_phon_min, 
                     col_end = n_phon_max) %>%
    cols_merge_range(col_begin = lv_min,
                     col_end = lv_max) %>%
    cols_label(list = "",
               n = md("*N*"),
               freq_mean = md("Mean ± SD"),
               freq_min = md("Range"),
               n_phon_mean = md("Mean ± SD"),
               n_phon_min = md("Range"),
               lv_mean = md("Mean ± SD"),
               lv_min = md("Range")) %>%
    tab_style(cell_text(style = "italic"),
              cells_column_labels(everything())) %>%
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners(everything())) %>%
    summary_rows(fns = list("Grand mean" = ~mean(.)),
                 columns = ends_with("_mean")) %>%
    summary_rows(fns = list("N" = ~sum(.)),
                 columns = n,
                 formatter = fmt_integer)
```

@@

```{r}
#| label: fig-items-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 9
#| fig-cap: "Distribution of Levenshtein similarities across items"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv) %>% 
    mutate(prop = n/sum(n),
           labels = paste0(
               n, " (",
               percent(prop, accuracy = 0.1), 
               ")")) %>% 
    ggplot(aes(x = lv, n)) +
    geom_col(fill = "grey60", colour = "white") +
    geom_text(aes(label = labels),
              position = position_nudge(y = 3), 
              size = 3.5) +
    labs(x = "Levenshtein distance", y = "Number of words") +
    theme(
        panel.grid.major.x = element_blank(),
        legend.position = "top"
    )
```

@@

```{r}
#| label: fig-items-frequency-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(freq, lv) %>% 
    ggplot(aes(freq, lv)) +
    geom_point(size = 2,
               shape = 1, 
               colour = "grey60",
               stroke = 1) +
    geom_smooth(method = "lm", 
                formula = "y ~ x", 
                colour ="black") +
    labs(y = "Levenshtein similarity", 
         x = "Lexical frequency (Zipf score)", 
         colour = "Language", 
         fill = "Language") +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r}
#| label: fig-items-nphon-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association between number of phonemes and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(n_phon, lv) %>% 
    ggplot(aes(n_phon, lv, size = n)) +
    geom_point(shape = 1,
               stroke = 1,
               colour = "grey60") +
    geom_smooth(method = "lm",
                formula = "y ~ x",
                size = 1,
                colour = "black") +
    labs(x = "# Phonemes", 
         y = "Levenshtein similarity", 
         colour = "Language", 
         fill = "Language") +
    scale_y_continuous(labels = percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank()) 
```

@@

thresholds %>%
    ggplot(aes(ymin = ymin, ymax = ymax,
               xmin = xmin, xmax = xmax,
               fill = .category)) +
    geom_rect(alpha = 0.5) +
    stat_dist_slab(
        data = dist_df, 
        aes(dist = "norm", arg1 = mean, arg2 = sd),
        position = "dodge", inherit.aes = FALSE,
        fill = NA, colour = "black"
    ) +
    geom_hline(yintercept = c(-1.5, 1.5), 
               linetype = "dotted",
               linewidth = 1) +
    geom_text(aes(label = .category, y = c(-3, 0, 3), x = 1),
              vjust = 1, hjust = 0.5) +
    annotate(
        geom = "text",
        label = c("Threshold~1~(p[1])", "Threshold~2~(p[2])"),
        y = c(-1.5, 2), parse = TRUE,
        x = 0.5, angle = 90, vjust = -0.5
    ) +
    labs(y = "Hypothetical response-generating latent distribution",
         x = "Response probability",
         fill = "Response type",
         colour = "Response type") +
    coord_flip() +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_x_continuous(labels = percent) +
    scale_y_continuous(limits = c(-4, 4)) +
    theme(panel.grid = element_blank(),
          legend.position = "none")
```

@@

```{r}
#| label: fig-prior-mean
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected prior predictions"
nd <- expand.grid(
    n_phon_std = 0,
    age_std = (10:36-mean(responses$age))/(sd(responses$age)),
    exposure_std = 0,
    lv_std = 0
)

m <- add_epred_draws(nd, 
                     model_fit_4_prior, 
                     ndraws = NULL,
                     re_formula = NA) %>% 
    filter(.category != "No") %>% 
    pivot_wider(names_from = ".category", values_from = ".epred") %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(
        c(Understands, `Understands and Says`),
        names_to = ".category",
        values_to = ".epred"
    ) %>% 
    mutate(
        .category = factor(
            .category,
            levels = c("Understands and Says", "Understands"),
            labels = c("Production\n(Understands and Says))", "Comprehension\n(Understands)")
        )
    )

m %>% 
    ggplot(aes(x = age_std, y = .epred)) +
    facet_wrap(~.category) +
    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
    stat_summary(fun = "mean", geom = "line", size = 0.75) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(responses$age))/(sd(responses$age)),
        labels = seq(10, 36, 4)
    ) +
    scale_fill_manual(
        values = rev(clrs), 
        labels = c("95%", "89%", "67%", "50%")
    ) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Credible interval", 
        fill = "Credible interval"
    ) +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank()
    )
```

@@

m <- add_predicted_draws(nd, model_fit_4_prior, ndraws = NULL, re_formula = NA)

m %>% 
    count(age_std, .prediction) %>% 
    pivot_wider(names_from = .prediction, values_from = n) %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(c(No, Understands, `Understands and Says`), names_to = ".prediction", values_to = "n") %>%
    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
    geom_col(position = position_fill()) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Category", 
        fill = "Category"
    ) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(responses$age))/(sd(responses$age)),
        labels = seq(10, 36, 4)
    ) +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

@@

### Model comparison

To test of goodness of fit of our model, we compared its predictive accuracy of this extended model--which included all predictors of interest--to four simpler models that included a subset of our predictors of interest: a base model that only included the main effects of $Age$, $Phonemes$, and $Exposure$ ($\mathcal{M}_{0}$), a model that also included the $Age \times Exposure$ interaction ($\mathcal{M}_{1}$), a model that also included the main effect of $Levenshtein$ ($\mathcal{M}_{2}$), a model that also included the $Age \times Levenshtein$ interaction ($\mathcal{M}_{3}$). Only the extended model included the $Age \times Exposure \times Levenshtein$ interaction. We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (*LOO*), a relative fit index, via the `loo` R package. LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set. Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model.

@@

#### Model 0

```{r}
print(model_fit_0$formula)
```

#### Model 1

```{r}
print(model_fit_1$formula)
```

#### Model 2

```{r}
print(model_fit_2$formula)
```

#### Model 3

```{r}
print(model_fit_3$formula)
```

#### Model 4

```{r}
print(model_fit_4$formula)
```
:::

@@

## Model selection

>>>>

---
title: "Cognate beginnings to bilingual lexical acquisition"
subtitle: "Lab notes"
short-title: "Cognate beginnings"
author:
  - name: Gonzalo Garcia-Castro
    corresponding: true
    orcid: 0000-0002-8553-4209
    email: gonzalo.garciadecastro@upf.edu
    affiliations:
      - id: cbc
        name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
        address: Ramon Trias Fargas 25-27
        city: Barcelona
        region: Spain
        postal-code: "08005"
  - name: Daniela S. Avila-Varela
    orcid: 0000-0002-3518-8117
    affiliations: 
      - ref: cbc
      - id: ul
        name: Universidade de Lisboa
        department: Facultade de Letras
        address: Alameda da Universidade
        city: Lisboa
        region: Portugal
        postal-code: "1600-214"
  - name: Ignacio Castillejo
    orcid: 0000-0001-7445-0416
    affiliations: 
      - id: uam 
        name: Universidad Autónoma de Madrid
        department: Departamento de Psicología
        address: Iván Pavlov, 6
        city: Madrid
        region: Spain
        postal-code: "28049"
  - name: Núria Sebastian-Galles
    orcid: 0000-0001-6938-2498
    affiliations: 
      - ref: cbc
date: "May 10th, 2023"
code-repo: "Access the code, data, and analysis at <https://github.com/gongcastro/trajectories>"
keywords:
  - lexical acquisition
  - vocabulary
  - bilingualism
  - item response theory
  - bayesian
csl: "../assets/apa7.csl"
bibliography: "../assets/references.bib"
link-citations: true
editor: source

format:
  html: default

abstract: |
    Language non-selectivity is prominent feature of bilingual lexical processing. An instance of such non-selectivity is embodied by cognateness, which recent studies have suggested to facilitate vocabulary acquisition in bilingual toddlers, who show larger vocabulary sizes when their languages share many cognates, and to be acquired at earlier ages than non-cognates. The specific mechanisms behind such facilitation are unclear. We present an account of early bilingual vocabulary growth in which cognateness interacts with lexical frequency and language exposure to facilitate the word acquisition. We evaluated this model against vocabulary data from 436 Catalan-Spanish bilinguals toddlers. We used Bayesian Explanatory Item Response Models to estimate participants’ probability of acquisition of 604 words, conditional to the cognate status and lexical frequency of the word-form, and the age and degree of exposure to each language of the toddler. We found cognateness facilitation effect only in low-frequency words. Overall, our findings support an interactive account of bilingual vocabulary acquisition in which the lexical representations in one language interact with the acquisition of words in the other language.

fig-dpi: 1000

---

```{r setup}
#| label: setup
#| echo: false
#| message: false
#| warning: false
# load objects -----------------------------------------------------------------

library(patchwork)
library(gt)
library(ggplot2)

targets::tar_config_set(store = here::here("_targets"),
                        script = here::here("_targets.R"))

targets::tar_load_globals()

targets::tar_load(c(bvq_data, childes, items, 
                    participants, responses))

targets::tar_load(c(model_fit, model_summary, model_epreds,
                    model_convergence, model_fit_prior, 
                    model_ppcs, posterior_doe_summary,
                    model_draws, model_summary,
                    posterior_doe_draws))

targets::tar_load(c(syllables_data, model_fit_syllables, post_syllables))

# process objects
items <- bvq_data$pool  |>  
    mutate(item = str_remove(item, "cat_|spa_")) |> 
    select(item, language, semantic_category, class) |> 
    right_join(items, by = c("item", "language"))

rope_interval <- c(lower = -0.1, upper = 0.1)

# set ggplot theme and colour palette ------------------------------------------

my_theme <- theme_minimal() +
    theme(panel.grid = element_line(colour = "grey",
                                    linetype = "dotted"),
          axis.line = element_line(colour = "black"),
          text = element_text(size = 12, colour = "black"),
          axis.text = element_text(colour = "black"))

theme_set(my_theme)

bayesplot::bayesplot_theme_set(my_theme + theme(panel.grid = element_blank()))

clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

set.seed(888)

options(ggplot2.ordinal.fill = clrs,
        ggplot2.ordinal.colour = clrs,
        ggplot2.discrete.fill = clrs,
        ggplot2.discrete.colour = clrs,
        ggplot2.continuous.fill = scale_color_gradient,
        ggplot2.continuous.colour = scale_color_gradient,
        knitr.graphics.error = FALSE) 
```

@@

```{r}
#| label: questionnaire
#| echo: false
#| message: false
#| warning: false
data("pool", package = "bvq")
n_categories <- distinct(pool, semantic_category) %>% nrow()
n_items <- length(unique(pool$item))
n_item_language <- count(pool, language) %>% 
    pull(n) %>%
    set_names(c("catalan", "spanish"))
```

@@

```{r apatb-participants}
#| label: tab-participants
#| echo: false
#| message: false
#| warning: false
#| tbl-cap: "Participant sample size by age and degree of exposure to Catalan. Information about language degree of exposure (DoE) was provided by families before filling the questionnaire. In this table, a 100\\% indicates that the participant was exclusively exposed to Catalan, and never to Spanish. A 0\\% indicates that the participant was not exposed to Catalan ever, and most of the time to Spanish. A 50\\% indicates that the participant was exposed to Catalan and Spanish approximately half of the time each. The 100\\%, 0\\%, and 50\\% would be traditionaly classified as Catalan monolingual, Spanish monolingual, and Catalan-Spanish bilingual, respectively. For illustration purposes this table, DoEs were binned into 25\\%-wide bins, and ages (in months) binned into 4 month-wide bins. Hyphens indicate that no participants from that specific combination of age and DoE filled the questionnaire."
participants |>
    mutate_at(vars(starts_with("doe_")), 
              (\(x) floor(x * 10) / 10)) |> 
    mutate(
        age = cut(floor(age),
                  c(10, 14, 18, 22, 26, 30, 34, 36),
                  include.lowest = TRUE),
        doe_catalan = cut(
            doe_catalan,
            c(0, 0.25, 0.5, 0.75, 1),
            labels = c("0-25%",
                       "25-50%", 
                       "50-75%",
                       "75-100%"),
            include.lowest = TRUE
        ) |>
            as.character()
    ) |>
    count(age, doe_catalan) |>
    arrange(desc(doe_catalan)) |>
    pivot_wider(names_from = age, values_from = n) |>
    gt() |>
    sub_missing(everything(), everything(), missing_text = "--") |>
    cols_label(doe_catalan = "Catalan exposure") |>
    tab_spanner("Age (months)", 2:7) |>
    tab_style(cell_text(align = "left"),
              list(cells_body(doe_catalan),
                   cells_column_labels(doe_catalan)))
```

@@

```{r}
#| label: tbl-participants-edu
#| tbl-cap: "Participant exposure to the language (DoE) and socio-economic status (SES)."
participants %>%
    mutate(id = as.character(id)) |> 
    left_join(bvq_data$logs,
              by = join_by(id, time, age, lp, doe_catalan, doe_spanish, edu_parent)) %>%
    mutate(
        edu_parent = edu_parent |> 
            as.factor() |> 
            fct_na_value_to_level(),
        doe_catalan = cut(doe_catalan,
                          breaks = seq(0, 1, 0.2),
                          include.lowest = TRUE)
    ) %>%
    count(doe_catalan, edu_parent) %>%
    right_join(
        expand_grid(
            doe_catalan = levels(.$doe_catalan),
            edu_parent = unique(bvq_data$logs$edu_parent)
        ),
        by = join_by(doe_catalan, edu_parent)
    ) %>%
    rename(total = n) %>%
    pivot_wider(names_from = edu_parent,
                values_from = total,
                values_fill = 0) %>%
    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>%
    rename(Total = total) %>%
    select(-"NA") %>%
    gt() |> 
    cols_label(doe_catalan = "DoE Catalan") %>%
    sub_missing(columns = everything(),
                missing_text = "--") %>%
    tab_spanner("Educational attainment", 
                Secondary:Primary) %>%
    tab_style(
        style = list(cell_borders(weight = px(2),
                                  sides = "left", 
                                  color = "grey50")),
        locations = cells_body(columns = 8,
                               rows = 1:5)
    ) %>%
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(columns = 2:7)
    ) %>%
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = c(1, 8))
    ) %>%
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(spanners = "Educational attainment")
    )
```

@@

items %>%
    unnest(list) %>%
    summarise(n = n(),
              across(c(freq, n_phon, lv),
                     lst(mean, sd, min, max)),
              .by = c(language, list)) |> 
    gt(groupname_col = "language", 
       rowname_col = "list") %>%
    fmt_number(-c(list, n),
               decimals = 1) %>%
    tab_spanner("Frequency (Zipf)",
                columns = starts_with("freq_")) %>%
    tab_spanner("Length (Phonemes)",
                columns = starts_with("n_phon_")) %>%
    tab_spanner("Levenshtein (cognateness)", 
                columns = starts_with("lv_")) %>%
    cols_merge_uncert(col_val = freq_mean,
                      col_uncert = freq_sd) %>%
    cols_merge_uncert(col_val = n_phon_mean,
                      col_uncert = n_phon_sd) %>%
    cols_merge_uncert(col_val = lv_mean, 
                      col_uncert = lv_sd) %>%
    cols_merge_range(col_begin = freq_min, 
                     col_end = freq_max) %>%
    cols_merge_range(col_begin = n_phon_min, 
                     col_end = n_phon_max) %>%
    cols_merge_range(col_begin = lv_min,
                     col_end = lv_max) %>%
    cols_label(list = "",
               n = md("*N*"),
               freq_mean = md("Mean ± SD"),
               freq_min = md("Range"),
               n_phon_mean = md("Mean ± SD"),
               n_phon_min = md("Range"),
               lv_mean = md("Mean ± SD"),
               lv_min = md("Range")) %>%
    tab_style(cell_text(style = "italic"),
              cells_column_labels(everything())) %>%
    tab_style(cell_text(weight = "bold"),
              cells_column_spanners(everything()))
```

@@

```{r}
#| label: fig-items-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 9
#| fig-cap: "Distribution of Levenshtein similarities across items"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv) %>% 
    mutate(prop = n/sum(n),
           labels = paste0(
               n, " (",
               scales::percent(prop, accuracy = 0.1), 
               ")")) %>% 
    ggplot(aes(x = lv, n)) +
    geom_col(fill = "grey60", colour = "white") +
    geom_text(aes(label = labels),
              position = position_nudge(y = 3), 
              size = 3.5) +
    labs(x = "Levenshtein distance", y = "Number of words") +
    theme(
        panel.grid.major.x = element_blank(),
        legend.position = "top"
    )
```

@@

```{r}
#| label: fig-items-frequency-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(freq, lv) %>% 
    ggplot(aes(freq, lv)) +
    geom_point(size = 2,
               shape = 1, 
               colour = "grey60",
               stroke = 1) +
    geom_smooth(method = "lm", 
                formula = "y ~ x", 
                colour ="black") +
    labs(y = "Levenshtein similarity", 
         x = "Lexical frequency (Zipf score)", 
         colour = "Language", 
         fill = "Language") +
    scale_y_continuous(labels = scales::percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r}
#| label: fig-items-nphon-lv
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association between number of phonemes and phonological similarity"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    drop_na(freq) %>% 
    count(n_phon, lv) %>% 
    ggplot(aes(n_phon, lv, size = n)) +
    geom_point(shape = 1,
               stroke = 1,
               colour = "grey60") +
    geom_smooth(method = "lm",
                formula = "y ~ x",
                size = 1,
                colour = "black") +
    labs(x = "# Phonemes", 
         y = "Levenshtein similarity", 
         colour = "Language", 
         fill = "Language") +
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(legend.position = "none",
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank()) 
```

@@

thresholds %>%
    ggplot(aes(ymin = ymin, ymax = ymax,
               xmin = xmin, xmax = xmax,
               fill = .category)) +
    geom_rect(alpha = 0.5) +
    stat_dist_slab(
        data = dist_df, 
        aes(dist = "norm", arg1 = mean, arg2 = sd),
        position = "dodge", inherit.aes = FALSE,
        fill = NA, colour = "black"
    ) +
    geom_hline(yintercept = c(-1.5, 1.5), 
               linetype = "dotted",
               linewidth = 1) +
    geom_text(aes(label = .category, y = c(-3, 0, 3), x = 1),
              vjust = 1, hjust = 0.5) +
    annotate(
        geom = "text",
        label = c("Threshold~1~(p[1])", "Threshold~2~(p[2])"),
        y = c(-1.5, 2), parse = TRUE,
        x = 0.5, angle = 90, vjust = -0.5
    ) +
    labs(y = "Hypothetical response-generating latent distribution",
         x = "Response probability",
         fill = "Response type",
         colour = "Response type") +
    coord_flip() +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_x_continuous(labels = scales::percent) +
    scale_y_continuous(limits = c(-4, 4)) +
    theme(panel.grid = element_blank(),
          legend.position = "none")
```

@@

```{r fig_prior-mean}
#| label: fig-prior-mean
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 3
#| out-width: 90%
#| fig-cap: "Expected prior-predicted mean."
nd <- expand.grid(n_phon_std = 0,
                  age_std = (10:36-mean(responses$age))/(sd(responses$age)),
                  exposure_std = 0,
                  lv_std = 0)

prior_epreds <- add_epred_draws(nd, 
                                model_fit_prior,
                                ndraws = 50, 
                                re_formula = NA) |> 
    filter(.category != "No") |> 
    pivot_wider(names_from = ".category", values_from = ".epred") |> 
    mutate(Understands = Understands + `Understands and Says`) |> 
    pivot_longer(c(Understands, `Understands and Says`),
                 names_to = ".category",
                 values_to = ".epred") |> 
    mutate(.category = factor(.category,
                              levels = c("Understands",
                                         "Understands and Says"),
                              labels = c("Comprehension\n(Understands)",
                                         "Production\n(Understands and Says)")))

prior_epreds |> 
    ggplot(aes(x = age_std, y = .epred)) +
    facet_wrap(~ .category) +
    geom_line(linewidth = 3/4,
              alpha = 1/2,
              aes(group = .draw),
              colour = "grey50") + 
    labs(x = "Age (months)", 
         y = "Probability of acquisition",
         colour = "Credible interval", 
         fill = "Credible interval") +
    scale_x_continuous(breaks = scale(seq(10, 36, 4), 
                                      mean(responses$age), 
                                      sd(responses$age))[, 1],
                       labels = seq(10, 36, 4)) +
    theme(legend.position = "top",
          axis.ticks.x = element_line(colour = "black"),
          panel.grid = element_blank())
```

@@

m <- add_predicted_draws(nd, model_fit_prior, ndraws = NULL, re_formula = NA)

m %>% 
    count(age_std, .prediction) %>% 
    pivot_wider(names_from = .prediction, values_from = n) %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(c(No, Understands, `Understands and Says`),
                 names_to = ".prediction", 
                 values_to = "n") %>%
    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
    geom_col(position = position_fill()) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Category", 
        fill = "Category"
    ) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(responses$age))/(sd(responses$age)),
        labels = seq(10, 36, 4)
    ) +
    scale_fill_manual(values = clrs[c(1, 4, 5)]) +
    scale_y_continuous(labels = scales::percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

@@

@@

```{r}
print(model_fit$formula)
```
:::

@@
